Models of VOR learning in MHC knockout mice

### 2013-02-12 Notes/VORmodel.tex
Discussion of OG multistate

\tableofcontents

@@

\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[height=2cm]{cascade.svg}}\label{fig:cascade_model}\hspace{0.5cm}
  \item\aligntop{\includegraphics[height=1.7cm]{multistate.svg}}\label{fig:multistate_model}\hspace{0.5cm}
  \item\aligntop{\includegraphics[height=1.7cm]{binary.svg}}\label{fig:binary_model}
 \end{myenuma}
 \end{center}
  \caption{Transition probabilities for different models.
  (\ref{fig:cascade_model}) In the cascade model, the transition probabilities decay geometrically with a parameter $x$ (see \cite{Fusi2005cascade}).
  (\ref{fig:multistate_model}) In the multistate model the transition probabilities for potentiation/depression are all equal and it is parameterised by these two values.
  (\ref{fig:binary_model}) The two-state model is parameterised by the two transition probabilities.} \label{fig:models}
\end{figure}

\begin{table}
 \begin{center}
  \begin{tabularn}{|l|c|c|c|c|c|c|}
    \cline{1-7}
    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
    Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
    \cline{1-7}
    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3  & 50 & 50  &\label{tr:cascade_short} \\
    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3  & 50 & 150 &\label{tr:cascade_long} \\
    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1  & 50 & 50  &\label{tr:multistate_weak} \\
    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.17 & 10 & 30  &\label{tr:multistate_med} \\
    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4  & 20 & 20  &\label{tr:multistate_strong} \\
    Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1  & 5  & 5   &\label{tr:binary} \\
    Pooled     & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
                                          & -0.1 & 70 & 70 &\label{tr:pooled_plenty}\\
    Pooled     & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
                                          & -0.1 & 70 & 70 &\label{tr:pooled_scarce}\\
    \cline{1-7}
%   \hline
%    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
%    & Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
%    \cline{1-7}
%    \label{tr:cascade_short}&
%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 50   \\
%    \label{tr:cascade_long}&
%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 150  \\
%    \label{tr:multistate_weak}&
%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1 & 50 & 50   \\
%    \label{tr:multistate_strong}&
%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4 & 20 & 20   \\
%    \label{tr:binary}&
%    Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1 & 5  & 5    \\
%    \label{tr:pooled_plenty}&
%    Pooled     & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
%                                          & -0.1 & 70 & 70 \\
%    \label{tr:pooled_scarce}&
%    Pooled     & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
%                                          & -0.1 & 70 & 70 \\
%    \cline{1-7}
  \end{tabularn}
 \end{center}
  \caption{Parameters used in simulations.} \label{tab:params}
\end{table}

@@

We will model the individual synapses with the two-state model.
Let $i$ be the number of them that are potentiated.
We will model the effect of resource depletion linearly with the potentiation/depression probabilities for the individual synapses:
%
\begin{equation}\label{eq:depletion}
  \begin{aligned}
    q\pot &= \frac{(P-i-1)q\lmax + i\,q\lmin}{P-1}, \qquad
    q\dep &= \frac{(i-1)q\lmax + (P-i)q\lmin}{P-1}.
  \end{aligned}
\end{equation}
%
At each plasticity event for the compound synapse, one of the individual synapses will be chosen randomly for update.
This effectively reduces the transition probabilities by $1/P$.

@@

We will try to gain some analytic insight to some of these models by looking at the slope of the learning curve at the start of gain-increase training.
This is proportional to the net-flux between the $\w=+1$ states and the $\w=-1$ states, measuring using the transition probabilities for gain-increase but the equilibrium distribution for either untrained or gain-decrease, assuming that pre-training lasts long enough to reach the equilibrium distribution for gain-decrease.

\newcommand{\resultsfig}[2]{\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=7cm]{#1_learn.eps}}\label{fig:#1_learn}
  \item\aligntop{\includegraphics[width=7cm]{#1_learnS.eps}}\label{fig:#1_learnS}
  \item\label{fig:#1_eq}\begin{myenumi}
                    \item\aligntop{\includegraphics[width=7cm]{#1_eq_WT.eps}}\label{fig:#1_eq_WT}
                    \item\aligntop{\includegraphics[width=7cm]{#1_eq_KO.eps}}\label{fig:#1_eq_KO}
                  \end{myenumi}
  \item\label{fig:#1_pr}\begin{myenumi}
                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_WT_nopre.eps}}\label{fig:#1_pr_WT_nopre}
                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_WT_pre.eps}}\label{fig:#1_pr_WT_pre}
                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_KO_nopre.eps}}\label{fig:#1_pr_KO_nopre}
                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_KO_pre.eps}}\label{fig:#1_pr_KO_pre}
                  \end{myenumi}
 \end{myenuma}
 \end{center}
  \caption[#2]{#2.
  Other parameters can be found in \autoref{tr:#1} of \autoref{tab:params}.
  Time is measured in units of $1/r$, the mean time between plasticity events.
  (\ref{fig:#1_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
  (\ref{fig:#1_learnS}) Learning curves restricted to gain-increase training.
  (\ref{fig:#1_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#1_eq_WT}) wild-type and (\ref{fig:#1_eq_KO}) \KO\ mutant.
  (\ref{fig:#1_pr}) Evolution of probability distributions for (\ref{fig:#1_pr_WT_nopre}\ref{fig:#1_pr_WT_pre}) wild-type and  (\ref{fig:#1_pr_KO_nopre}\ref{fig:#1_pr_KO_pre}) \KO\ mutant without (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_KO_nopre}) and with (\ref{fig:#1_pr_WT_pre},\ref{fig:#1_pr_KO_pre}) pre-training. } \label{fig:#1}
\end{figure}}

@@

The results of simulations of the cascade model can be seen in \autoref{fig:cascade_short} and \autoref{fig:cascade_long}.

@@

First, consider the wild-type, for which $q\pot=q\dep=q$.
Without pre-training:
%
\begin{equation}\label{eq:multiWTnopre}
  \Phi = -\frac{2\Delta fq}{M},
\end{equation}
%
where it's worth remembering that $\Delta f<0$.
With pre-training:
%
\begin{equation}\label{eq:multiWTpre}
\begin{aligned}
  \Phi &= 16(\Delta f)^2q \, \frac{(1-2\Delta f)^{M/2-1} (1+2\Delta f)^{M/2-1}}
          {(1-2\Delta f)^M - (1+2\Delta f)^M} \\
       &= -\frac{4\Delta fq}{M} + \CO(\Delta f)^2.
\end{aligned}
\end{equation}
%
So, we see that pre-training will speed up learning when $\Delta f$ is small (in absolute value), as seen in \autoref{fig:multistate_weak}\ref{fig:multistate_weak_learnS}.
On the other hand, if $\Delta f$ is close to $-\half$, pre-training will initially slow down learning a lot, as seen in \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}.

@@

With pre-training:
%
\begin{equation}\label{eq:multiKOpre}
\begin{aligned}
  \Phi &= -4\Delta f q \, \frac{(1+2\Delta f) - \beta(1-2\Delta f)}
          {(1+2\Delta f)^M - \beta^M(1-2\Delta f)^M}   \,
          \beta^{M/2-1}(1-2\Delta f)^{M/2-1} (1+2\Delta f)^{M/2-1} \\
       &= -4\Delta f q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M} + \CO(\Delta f)^2.
\end{aligned}
\end{equation}
%
Once again, we see that pre-training will speed up learning when $\Delta f$ is small, whereas, if $\Delta f$ is close to $-\half$, pre-training will initially slow down learning.

@@

Now consider what would happen if we did not treat gain-increase and gain-decrease symmetrically.
Let us set $f\pot\dec = f\pot\norm - \overline{\Delta f}$.
Then, \eqref{eq:multiKOpre} becomes
%
\begin{equation}\label{eq:multiKOdiffpre}
\begin{aligned}
  \Phi &= -2\prn{\Delta f + \overline{\Delta f}} q \, \frac{(1+2\overline{\Delta f}) - \beta(1-2\overline{\Delta f})}
          {(1+2\overline{\Delta f})^M - \beta^M(1-2\overline{\Delta f})^M}   \,
          \beta^{M/2-1}(1-2\overline{\Delta f})^{M/2-1} (1+2\overline{\Delta f})^{M/2-1} \\
       &= -2\prn{\Delta f + \overline{\Delta f}} q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M} + \CO(\Delta f,\overline{\Delta f})^2.
\end{aligned}
\end{equation}
%
We can then look at the wild-type by setting $\beta=1$:
%
\begin{equation}\label{eq:multiWTdiffpre}
\begin{aligned}
  \Phi &= -8\,\overline{\Delta f}\prn{\Delta f + \overline{\Delta f}} q \,
          \frac{(1-2\overline{\Delta f})^{M/2-1} (1+2\overline{\Delta f})^{M/2-1} }
          {(1+2\overline{\Delta f})^M - (1-2\overline{\Delta f})^M}   \\
       &= -\frac{2\prn{\Delta f + \overline{\Delta f}} q}{M} + \CO(\Delta f,\overline{\Delta f})^2.
\end{aligned}
\end{equation}
%

We can see from these formulae that, if we want pre-training to impair learning in wild-type, but enhance it in the mutant, the important thing is to have \emph{strong pre-training}, not strong training.
Thus, removing the symmetry between gain-increase and gain-decrease will not help with the unrealistic feature of this model -- the fact that pre-training proceeds much faster in the mutant than wild-type.

\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=7cm]{multistate_betastar.eps}}\label{fig:multistate_betastar}
  \item\aligntop{\includegraphics[width=7cm]{multistate_deltafstar.eps}}\label{fig:multistate_deltafstar}
 \end{myenuma}
 \end{center}
  \caption{The functions (\ref{fig:multistate_betastar}) $\beta^*(M)$, which describes when the mutant has impaired learning, and (\ref{fig:multistate_deltafstar}) $\Delta f^*(\beta,M)$ for $M=10$, which describes when pre-training enhances learning.}\label{fig:multistate_star}
\end{figure}

@@

In this section, we will consider the multistate model as originally defined in \cite{amit1994learning}, \ie with linearly varying synaptic weight:
%
\begin{equation}\label{eq:multistateLinWeight}
  \w_i = \frac{2i-M-1}{M-1}.
\end{equation}
%
The equilibrium distribution, \eqref{eq:mutltieq}, still applies.
However, now the rate of change of our learning metric, \eqref{eq:learning}, will be proportional to the sum of the net fluxes between adjacent states:
%
\begin{equation}\label{eq:multiLinFlux}
  \begin{aligned}
    \Phi &= \sum_{i=1}^{M-1} \eq_{i+1} f'{}\dep q\dep - \eq_i f'{}\pot q\pot
         &= \sum_{i=1}^{M-1} \eq_i (\alpha-\alpha') f'{}\dep q\dep \\
         &= \frac{1-\alpha^{M-1}}{1-\alpha^M} \, (\alpha-\alpha') f'{}\dep q\dep.
  \end{aligned}
\end{equation}
%

@@

Now, consider the mutant, for which $q\pot=\beta q\dep=q$, $\beta<1$.
Without pre-training:
%
\begin{equation}\label{eq:multiLinKOnopre}
  \Phi = -2\Delta f q\,\frac{1-\beta^{M-1}}{1-\beta^M}.
\end{equation}
%
With pre-training:
%
\begin{equation}\label{eq:multiLinKOpre}
\begin{aligned}
  \Phi &= -4\Delta f q \, \frac{(1+2\Delta f)^{M-1} - \beta^{M-1}(1-2\Delta f)^{M-1}}
          {(1+2\Delta f)^M - \beta^M(1-2\Delta f)^M} \\
       &= -4\Delta f q\,\frac{1-\beta^{M-1}}{1-\beta^M} + \CO(\Delta f)^2.
\end{aligned}
\end{equation}
%

@@

\resultsfig{pooled_plenty}{Simulation results for the pooled resource model with light resource depletion ($q\lmin/q\lmax=0.75$)}

\resultsfig{pooled_scarce}{Simulation results for the pooled resource model with heavy resource depletion ($q\lmin/q\lmax=0.25$)}

The results of simulations of the pooled resource model can be seen in \autoref{fig:pooled_plenty} and \autoref{fig:pooled_scarce}.

>>>>

\tableofcontents
\listoffigures

@@

\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[height=2cm]{cascade.svg}}\label{fig:cascade_model}\hspace{0.5cm}
  \item\aligntop{\includegraphics[height=1.7cm]{multistate.svg}}\label{fig:multistate_model}\hspace{0.5cm}
  \item\aligntop{\includegraphics[height=1.7cm]{binary.svg}}\label{fig:binary_model}
 \end{myenuma}
 \end{center}
  \caption[Transition probabilities for different models]{Transition probabilities for different models.
  (\ref{fig:cascade_model}) In the cascade model, the transition probabilities decay geometrically with a parameter $x$ (see \cite{Fusi2005cascade}).
  (\ref{fig:multistate_model}) In the multistate model the transition probabilities for potentiation/depression are all equal and it is parameterised by these two values.
  (\ref{fig:binary_model}) The two-state model is parameterised by the two transition probabilities.} \label{fig:models}
\end{figure}

\begin{table}
 \begin{center}
  \begin{tabularn}{|l|c|c|c|c|c|c|}
    \cline{1-7}
    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
    Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
    \cline{1-7}
    Cascade       & 10 & $x=0.25$ & $x=0.33$ & -0.3  & 50 & 50  &\label{tr:cascade_short} \\
    Cascade       & 10 & $x=0.25$ & $x=0.33$ & -0.3  & 50 & 150 &\label{tr:cascade_long} \\
    Multistate    & 10 & $q=0.6$  & $q=0.8$  & -0.1  & 50 & 50  &\label{tr:multistate_weak} \\
    Multistate    & 10 & $q=0.6$  & $q=0.8$  & -0.17 & 10 & 30  &\label{tr:multistate_med} \\
    Multistate    & 10 & $q=0.6$  & $q=0.8$  & -0.4  & 20 & 20  &\label{tr:multistate_strong} \\
    OG Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.3  & 20 & 20  &\label{tr:multistate_lin} \\
    Two-state     & 2  & $q=0.4$  & $q=0.8$  & -0.1  & 5  & 5   &\label{tr:binary} \\
    Pooled        & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
                                          & -0.1 & 70 & 70 &\label{tr:pooled_plenty}\\
    Pooled        & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
                                          & -0.1 & 70 & 70 &\label{tr:pooled_scarce}\\
    \cline{1-7}
%   \hline
%    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
%    & Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
%    \cline{1-7}
%    \label{tr:cascade_short}&
%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 50   \\
%    \label{tr:cascade_long}&
%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 150  \\
%    \label{tr:multistate_weak}&
%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1 & 50 & 50   \\
%    \label{tr:multistate_strong}&
%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4 & 20 & 20   \\
%    \label{tr:binary}&
%    Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1 & 5  & 5    \\
%    \label{tr:pooled_plenty}&
%    Pooled     & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
%                                          & -0.1 & 70 & 70 \\
%    \label{tr:pooled_scarce}&
%    Pooled     & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
%                                          & -0.1 & 70 & 70 \\
%    \cline{1-7}
  \end{tabularn}
 \end{center}
  \caption{Parameters used in simulations.} \label{tab:params}
\end{table}

@@

We will model the individual synapses with the two-state model.
Let $i=0\ldots P$ be the number of them that are potentiated.
We will model the effect of resource depletion linearly with the potentiation/depression probabilities for the individual synapses:
%
\begin{equation}\label{eq:depletion}
  \begin{aligned}
    q\pot &= \frac{(P-i-1)q\lmax + i\,q\lmin}{P-1}, \quad& i &= 0 \ldots P-1,\\
    q\dep &= \frac{(i-1)q\lmax + (P-i)q\lmin}{P-1}, & i &= 1 \ldots P.
  \end{aligned}
\end{equation}
%
At each plasticity event for the compound synapse, one of the individual synapses will be chosen randomly for update.
This effectively reduces the transition probabilities by $1/P$.

@@

We will try to gain some analytic insight to some of these models by looking at the slope of the learning curve at the start of gain-increase training.
This is proportional to the net-flux from the states with strong synaptic weight to the weaker states, measuring using the transition probabilities for gain-increase but the equilibrium distribution for either untrained or gain-decrease, assuming that pre-training lasts long enough to reach the equilibrium distribution for gain-decrease.
This should be multiplied by the difference in synaptic weights 
But, as this quantity is constant within a model and doesn't change when comparing wild-type and mutant or with/without pre-training, it can be ignored.

\newcommand{\resultsfig}[2]{\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=7cm]{#1_learn.eps}}\label{fig:#1_learn}
  \item\aligntop{\includegraphics[width=7cm]{#1_learnS.eps}}\label{fig:#1_learnS}
  \item\label{fig:#1_eq}\begin{myenumi}
                    \item\aligntop{\includegraphics[width=7cm]{#1_eq_WT.eps}}\label{fig:#1_eq_WT}
                    \item\aligntop{\includegraphics[width=7cm]{#1_eq_KO.eps}}\label{fig:#1_eq_KO}
                  \end{myenumi}
  \item\label{fig:#1_pr}\begin{myenumi}
                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_WT_nopre.eps}}\label{fig:#1_pr_WT_nopre}
                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_WT_pre.eps}}\label{fig:#1_pr_WT_pre}
                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_KO_nopre.eps}}\label{fig:#1_pr_KO_nopre}
                    \item\aligntop{\includegraphics[width=3cm]{#1_pr_KO_pre.eps}}\label{fig:#1_pr_KO_pre}
                  \end{myenumi}
 \end{myenuma}
 \end{center}
  \caption[#2]{#2.
  Other parameters can be found in \autoref{tr:#1} of \autoref{tab:params}.
  Time is measured in units of $1/r$, the mean time between plasticity events.
  (\ref{fig:#1_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
  (\ref{fig:#1_learnS}) Learning curves restricted to gain-increase training.
  (\ref{fig:#1_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#1_eq_WT}) wild-type and (\ref{fig:#1_eq_KO}) \KO\ mutant,
  with states of weak synaptic strength to the left and strong states to the right.
  (\ref{fig:#1_pr}) Evolution of probability distributions for (\ref{fig:#1_pr_WT_nopre}\ref{fig:#1_pr_WT_pre}) wild-type and  (\ref{fig:#1_pr_KO_nopre}\ref{fig:#1_pr_KO_pre}) \KO\ mutant without (\ref{fig:#1_pr_WT_nopre},\ref{fig:#1_pr_KO_nopre}) and with (\ref{fig:#1_pr_WT_pre},\ref{fig:#1_pr_KO_pre}) pre-training,
  with states of weak synaptic strength at the top and strong states underneath. } \label{fig:#1}
\end{figure}}

@@

The results of simulations of the cascade model can be seen in \autoref{fig:cascade_short} and \autoref{fig:cascade_long}.
It is difficult to study this model analytically.
However, the numerical results can helps us understand it qualitatively.

@@

First, consider the wild-type, for which $q\pot=q\dep=q$.
Without pre-training:
%
\begin{equation}\label{eq:multiWTnopre}
  \Phi = -\frac{2\Delta fq}{M},
\end{equation}
%
where it's worth remembering that $\Delta f<0$.
With pre-training:
%
\begin{equation}\label{eq:multiWTpre}
\begin{aligned}
  \Phi &= 16(\Delta f)^2q \, \frac{(1-2\Delta f)^{M/2-1} (1+2\Delta f)^{M/2-1}}
          {(1-2\Delta f)^M - (1+2\Delta f)^M} \\
       &= -\frac{4\Delta fq}{M} + \CO(\Delta f)^2.
\end{aligned}
\end{equation}
%
So, we see that pre-training will speed up learning when $\Delta f$ is small (in absolute value), as seen in \autoref{fig:multistate_weak}\ref{fig:multistate_weak_learnS}.
On the other hand, if $\Delta f$ is close to $-\half$, pre-training will initially slow down learning a lot due to the factor of $(1+2\Delta f)^{M/2-1}$, as seen in \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}.

@@

With pre-training:
%
\begin{equation}\label{eq:multiKOpre}
\begin{aligned}
  \Phi &= -4\Delta f q \, \frac{(1+2\Delta f) - \beta(1-2\Delta f)}
          {(1+2\Delta f)^M - \beta^M(1-2\Delta f)^M}   \,
          \brk{\beta (1-2\Delta f) (1+2\Delta f)}^{M/2-1} \\
       &= -4\Delta f q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M} + \CO(\Delta f)^2.
\end{aligned}
\end{equation}
%
Once again, we see that pre-training will speed up learning when $\Delta f$ is small, whereas, if $\Delta f$ is close to $-\half$, pre-training will initially slow down learning.

@@

Now consider what would happen if we did not treat gain-increase and gain-decrease symmetrically.
Let us set $f\pot\dec = f\pot\norm - \overline{\Delta f}$.
Then, \eqref{eq:multiKOpre} becomes
%
\begin{equation}\label{eq:multiKOdiffpre}
\begin{aligned}
  \Phi &= -2\prn{\Delta f + \overline{\Delta f}} q \, \frac{(1+2\overline{\Delta f}) - \beta(1-2\overline{\Delta f})}
          {(1+2\overline{\Delta f})^M - \beta^M(1-2\overline{\Delta f})^M}   \,
          \brk{\beta (1-2\overline{\Delta f}) (1+2\overline{\Delta f})}^{M/2-1} \\
       &= -2\prn{\Delta f + \overline{\Delta f}} q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M} + \CO(\Delta f,\overline{\Delta f})^2.
\end{aligned}
\end{equation}
%
We can then look at the wild-type by setting $\beta=1$:
%
\begin{equation}\label{eq:multiWTdiffpre}
\begin{aligned}
  \Phi &= -8\,\overline{\Delta f}\prn{\Delta f + \overline{\Delta f}} q \,
          \frac{(1-2\overline{\Delta f})^{M/2-1} (1+2\overline{\Delta f})^{M/2-1} }
          {(1+2\overline{\Delta f})^M - (1-2\overline{\Delta f})^M}   \\
       &= -\frac{2\prn{\Delta f + \overline{\Delta f}} q}{M} + \CO(\Delta f,\overline{\Delta f})^2.
\end{aligned}
\end{equation}
%

We can see from these formulae that, if we want pre-training to impair learning in wild-type, but enhance it in the mutant, the important thing is to have \emph{strong pre-training}, not strong training.
Thus, removing the symmetry between gain-increase and gain-decrease will not help with the most unrealistic feature of this model -- the fact that pre-training proceeds much faster in the mutant than wild-type.

\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=7cm]{multistate_betastar.eps}}\label{fig:multistate_betastar}
  \item\aligntop{\includegraphics[width=7cm]{multistate_deltafstar.eps}}\label{fig:multistate_deltafstar}
 \end{myenuma}
 \end{center}
  \caption[The functions $\beta^*(M)$ and $\Delta f^*(\beta,M)$]{The functions (\ref{fig:multistate_betastar}) $\beta^*(M)$, which describes when the mutant has impaired learning, and (\ref{fig:multistate_deltafstar}) $\Delta f^*(\beta,M)$ for $M=10$, which describes when pre-training enhances learning.}\label{fig:multistate_star}
\end{figure}

@@

\resultsfig{multistate_lin}{Simulation results for multistate model with linear weights}%\label{fig:multistate_lin}

In this section, we will consider the multistate model as originally defined in \cite{amit1994learning}, \ie with linearly varying synaptic weight:
%
\begin{equation}\label{eq:multistateLinWeight}
  \w_i = \frac{2i-M-1}{M-1}.
\end{equation}
%
The numerical results can be seen in \autoref{fig:multistate_lin}, but we can get some analytic insight into this model as well.

The equilibrium distribution, \eqref{eq:mutltieq}, still applies.
However, now the rate of change of our learning metric, \eqref{eq:learning}, will be proportional to the sum of the net fluxes between adjacent states:
%
\begin{equation}\label{eq:multiLinFlux}
  \begin{aligned}
    \Phi &= \sum_{i=1}^{M-1} \eq_{i+1} f'{}\dep q\dep - \eq_i f'{}\pot q\pot
         &= \sum_{i=1}^{M-1} \eq_i (\alpha-\alpha') f'{}\dep q\dep \\
         &= \frac{1-\alpha^{M-1}}{1-\alpha^M} \, (\alpha-\alpha') f'{}\dep q\dep.
  \end{aligned}
\end{equation}
%

@@

Now, consider the mutant, for which $q\pot=\beta q\dep=q$, $\beta<1$.
Without pre-training:
%
\begin{equation}\label{eq:multiLinKOnopre}
  \Phi = -2\Delta f q\,\frac{1-\beta^{M-1}}{1-\beta^M}.
\end{equation}
%
This decreases monotonically in the interval $\beta\in[0,1]$, therefore this will always be greater than \eqref{eq:multiLinWTnopre}.
This means that the mutant will initially learn faster than the wild-type.
However, as the wild-type will eventually catch up, and this can happen very quickly, as seen in \autoref{fig:multistate_lin}\ref{fig:multistate_lin_learn},\ref{fig:multistate_lin_learnS}

With pre-training:
%
\begin{equation}\label{eq:multiLinKOpre}
\begin{aligned}
  \Phi &= -4\Delta f q \, \frac{(1+2\Delta f)^{M-1} - \beta^{M-1}(1-2\Delta f)^{M-1}}
          {(1+2\Delta f)^M - \beta^M(1-2\Delta f)^M} \\
       &= -4\Delta f q\,\frac{1-\beta^{M-1}}{1-\beta^M} + \CO(\Delta f)^2.
\end{aligned}
\end{equation}
%
The ratio of this to \eqref{eq:multiLinKOnopre} takes its minimum value at the lowest end of the interval $\Delta f \in \brk{-\half,0}$, where it takes the value
%
\begin{equation}\label{eq:multiLinprenopre}
  \frac{\Phi_{\text{with pre}}}{\Phi_{\text{without pre}}} = \frac{1-\beta^M}{\beta-\beta^M}
   \longrightarrow \frac{M}{M-1} \quad \text{as} \quad \beta\to1,
\end{equation}
%
which is always greater than 1.
Therefore, pre-training will always enhance learning, for both mutant and wild-type.

@@

\resultsfig{pooled_plenty}{Simulation results for the pooled resource model with light depletion ($q\lmin/q\lmax=0.75$)}

\resultsfig{pooled_scarce}{Simulation results for the pooled resource model with heavy depletion ($q\lmin/q\lmax=0.25$)}

The results of simulations of the pooled resource model can be seen in \autoref{fig:pooled_plenty} and \autoref{fig:pooled_scarce}.
It is difficult to study this model analytically.
However, the numerical results can helps us understand it qualitatively.
